Learning to Synthesize Rules for Network Intrusion Detectors

### 2020-01-18 paper/paper.tex
pass on 4.3.4

\subsubsection{Frequency of content modifiers.} As previously
mentioned, \percRulesWithContent\ of Suricata rules involve content
options. Furthermore, we observed---by analyzing the dataset of
existing rules (see Section~\ref{sec:dataset-benign})---that often
these rules indicate what part of the message the content string
should be checked against.\Mar{Lucas, vc. tem dados para dar suporte a
  isto? ou seja, quantas options do tipo ``content'' possuem
  modificadores.} \Luc{das 26k regras que utilizam content, 13k delas utilizam modificadores http q sao os q nós consideramos.} For example, the
creator of the rule may want to indicate that a specific string be
present in the header of an HTTP request message. Content
\emph{modifiers} serve this purpose. For example, the modifier
\CodeIn{http\_header} in the option \CodeIn{content: ``keep-alive'';
  http\_header} states the requirement that the string ``keep-alive''
be present in the header of the message. (A keep-alive connection
attribute indicates that a single TCP connection should remain open
for multiple HTTP requests/responses.)
Figure~\ref{fig:http-header-example}

\begin{figure}[h!]
\centering
\includegraphics[scale=0.5]{figs/http_header-example.png}
\caption{\CodeIn{http\_header} example.}
\label{fig:http-header-example}
\end{figure}

%% Suricata supports content modifiers for tcp and http. \Luc{Eh possivel utilizar os modificadores em qualquer protocolo. Nós consideramos apenas tcp e http pois são os protocolos que possuem payloads legiveis, assim conseguimos saber melhor como tokenizar o conteudo. O payload de pacotes de outros protocolos, em geral, são um conjunto de bytes sem significado para nós. Existem diversos outros modificadores q não estamos considerando e q tambem permitem procurar numa determinada posição para os casos em q o conteudo do pacote n eh legivel, não citei pq não suportamos isso. Se necessário posso explicar.} Given that
%% content options often use modifiers\Mar{assumindo que eh
%%   verdade. confirmar estat. solicitada acima...},

\tname\ uses the frequency of content modifiers to discriminate rules
more or less likely to reflect actual rules. For example, historical
data indicates that content options that use the
\CodeIn{http\_connection} modifier are much less likely than content
options that use the modifier
\CodeIn{http\_uri}. Figure~\ref{fig:distribution-content_modifiers}
shows the histogram of content modifiers for http. Although content
modifiers are applicable to rules of any protocol supported by
Suricata, currently, \tname\ only supports tcp and http. These are the
protocols involving ~80\% of the rules from our dataset.

>>>>

\subsubsection{Frequency of content modifiers.} As previously
mentioned, \percRulesWithContent\ (=26K/\numrulessuri) of Suricata
rules involve content options. Furthermore, of the 26K rules that use
content options, 13K use HTTP modifiers\Mar{@Lucas, pode dar o numero
  desconsiderando HTTP? Esta parte eh generica.}. Content modifiers
serve to indicate what part of a network packet the string mentioned
in the content option should appear. For example, the modifier
\CodeIn{http\_header} in the option \CodeIn{content: ``keep-alive'';
  http\_header} states the requirement that the string ``keep-alive''
be present in the header of the message. (A keep-alive connection
attribute indicates that a single TCP connection should remain open
for multiple HTTP requests/responses.)
Figure~\ref{fig:http-header-example} illustrates possible content
options used to match against certain strings that appear in an HTTP
message. A check (resp., cross) mark indicates a match (resp.,
unmatch).

\begin{figure}[t!]
\centering
\includegraphics[scale=0.5]{figs/http_header-example.png}
\caption{\CodeIn{http\_header} example.}
\label{fig:http-header-example}
\end{figure}

%% we observed---by analyzing
%% the dataset of existing rules (see
%% Section~\ref{sec:dataset-benign})---that these rules often indicate
%% what part of the message the content string should be checked
%% against. For example,
%% For example, the
%% creator of the rule may want to indicate that a specific string be
%% present in the header of an HTTP request message. Content
%% \emph{modifiers} serve this purpose.
%% Suricata supports content modifiers for tcp and http. \Luc{Eh possivel utilizar os modificadores em qualquer protocolo. Nós consideramos apenas tcp e http pois são os protocolos que possuem payloads legiveis, assim conseguimos saber melhor como tokenizar o conteudo. O payload de pacotes de outros protocolos, em geral, são um conjunto de bytes sem significado para nós. Existem diversos outros modificadores q não estamos considerando e q tambem permitem procurar numa determinada posição para os casos em q o conteudo do pacote n eh legivel, não citei pq não suportamos isso. Se necessário posso explicar.} Given that
%% content options often use modifiers\Mar{assumindo que eh
%%   verdade. confirmar estat. solicitada acima...},

\tname{} leverages the fact that these modifiers are commonly used to
define rules as to better rank the rules it synthesizes. More
precisely, it builds on the observation that strings acessible in
positions associated with content modifiers (\eg{}, GET, Host,
Connection) are more likely to produce useful rules. \tname\ uses the
frequency of content modifiers to discriminate rules that are more or
less likely to reflect actual rules. For example, historical data
indicates that content options that use the \CodeIn{http\_connection}
modifier are much less likely than content options that use the
modifier
\CodeIn{http\_uri}. Figure~\ref{fig:distribution-content_modifiers}
shows the histogram of content modifiers for http. Although content
modifiers are applicable to rules of any protocol supported by
Suricata, currently, \tname\ only supports tcp and http. These are the
protocols involving ~80\% of the rules from our dataset.
